### 2017-08-14 sm.tex
SM update and other fixup

\documentclass[aps,secnumarabic,amsmath,amssymb]{revtex4}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{color}
\usepackage{graphics}
\usepackage[pdftex]{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[colorlinks=true]{hyperref}
\usepackage{listings}

\newcommand{\ud}{\mathrm{d}}
\newcommand{\ue}{\mathrm{e}}
\newcommand{\ui}{\mathrm{i}}
\newcommand{\res}{\mathrm{Res}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\dsum}{\displaystyle\sum}
\newcommand{\dprod}{\displaystyle\prod}
\newcommand{\dlim}{\dispqlaystyle\lim}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\fsno}[1]{{\!\not\!{#1}}}
\newcommand{\texp}[2]{\ensuremath{{#1}\times10^{#2}}}
\newcommand{\dexp}[2]{\ensuremath{{#1}\cdot10^{#2}}}
\newcommand{\eval}[2]{{\left.{#1}\right|_{#2}}}
\newcommand{\paren}[1]{{\left({#1}\right)}}
\newcommand{\lparen}[1]{{\left({#1}\right.}}
\newcommand{\rparen}[1]{{\left.{#1}\right)}}
\newcommand{\abs}[1]{{\left|{#1}\right|}}
\newcommand{\sqr}[1]{{\left[{#1}\right]}}
\newcommand{\crly}[1]{{\left\{{#1}\right\}}}
\newcommand{\angl}[1]{{\left\langle{#1}\right\rangle}}
\newcommand{\tpdiff}[4][{}]{{\paren{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}_{#4}}}
\newcommand{\tpsdiff}[4][{}]{{\paren{\frac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pdiff}[3][{}]{{\frac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\diff}[3][{}]{{\frac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psdiff}[3][{}]{{\frac{\partial^{#1}}{\partial {#3}{}^{#1}} {#2}}}
\newcommand{\sdiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\newcommand{\tpddiff}[4][{}]{{\left(\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}\right)_{#4}}}
\newcommand{\tpsddiff}[4][{}]{{\paren{\dfrac{\partial^{#1}}{\partial {#3}{}^{#1}}{#2}}_{#4}}}
\newcommand{\pddiff}[3][{}]{{\dfrac{\partial^{#1} {#2}}{\partial {#3}{}^{#1}}}}
\newcommand{\ddiff}[3][{}]{{\dfrac{\ud^{#1} {#2}}{\ud {#3}{}^{#1}}}}
\newcommand{\psddiff}[3][{}]{{\frac{\partial^{#1}}{\partial{}^{#1} {#3}} {#2}}}
\newcommand{\sddiff}[3][{}]{{\frac{\ud^{#1}}{\ud {#3}{}^{#1}} {#2}}}
\newcommand{\eff}{ef\! f}

\begin{document}
\title{Motional Ground-State Cooling Outside the Lamb-Dicke Regime -- Supplemental material}
\author{Y. Yu}
\email{yichaoyu@g.harvard.edu}
\author{N. R. Hutzler}
\altaffiliation{Present address: California Institute of Technology, Division of Physics, Mathematics, and Astronomy. Pasadena, CA, 91125}
\author{J. T. Zhang}
\author{L. R. Liu}
\author{T. Rosenband}
\author{K.-K. Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

\date{\today}

\maketitle

\section{Raman Sideband Cooling Pulse Sequence}

\subsection{Optical pumping parameters}
The optical pumping is done with a $\sigma^-$ polarized beam aligned with the bias magnetic field.
There are two frequencies in the beam that are on resonance with
the $|F=2,m_F=-1\rangle$ to $|F'=2,m_{F'}=-2\rangle$ D1 transition
and the $|F=1,m_F=-1\rangle$ to $|F'=2,m_{F'}=-2\rangle$ D2 transition respectively.
All optical pumping pulses have the same length of $40\mu$s
with a $|F=1,m_F=-1\rangle$ scattering rate of $0.14$ MHz
and a $|F=2,m_F=-1\rangle$ scattering rate of $0.39$ MHz.

\subsection{Raman pulse parameters}
Each of the Raman pulse in the cooling sequence is followed immediately by an optical pumping
pulse with the parameters given above.
The parameters for the Raman pulses are different between different groups and are listed below.
The full Rabi frequency is the two-photon Rabi frequency $\Omega$ excluding the matrix element
from wavefunction overlap.
For a Raman pulse with a power ramp,
the full Rabi frequency gives the arithmetics mean over the duration of the pulse.
The power ramp field gives, if any,
the beam (in terms of their corresponding $F$ number)
that ramps with a Blackman profile during the pulse~\cite{Kasevich1992}.
$\delta'$ is the two photon detuning relative to the zero field hyperfine splitting
of $1.7716261288(10)$GHz.
All of the relative detunings are negetive indicating that the frequency difference is smaller
than the zero field hyperfine splitting.
A more negative number corresponds to lower order cooling sidebands
and a more negative number corresponds to higher order ones.
$\delta'_0$ the relative detuning for the carrier with the specific powers in the Raman beams.
\subsubsection{Group 1}
This group is repeated 5 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $x$&-2&$-17.6375$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $y$&-2&$-17.2825$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $x$&-1&$-18.03$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $y$&-1&$-17.875$&$-18.47$&21&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 2}
This group is repeated 5 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-8&$-17.966$&$-18.506$&80&$2\pi\times14$&\\\hline
    $x$&-2&$-17.6375$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $z$&-7&$-18.0335$&$-18.506$&80&$2\pi\times14$&\\\hline
    $y$&-2&$-17.2825$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-8&$-17.966$&$-18.506$&80&$2\pi\times14$&\\\hline
    $x$&-1&$-18.03$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $z$&-7&$-18.0335$&$-18.506$&80&$2\pi\times14$&\\\hline
    $y$&-1&$-17.875$&$-18.47$&21&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 3}
This group is repeated 6 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-7&$-18.0335$&$-18.506$&80&$2\pi\times14$&\\\hline
    $x$&-2&$-17.6375$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $z$&-6&$-18.101$&$-18.506$&80&$2\pi\times14$&\\\hline
    $y$&-2&$-17.2825$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-7&$-18.0335$&$-18.506$&80&$2\pi\times14$&\\\hline
    $x$&-1&$-18.03$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $z$&-6&$-18.101$&$-18.506$&80&$2\pi\times14$&\\\hline
    $y$&-1&$-17.875$&$-18.47$&21&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 4}
This group is repeated 7 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-6&$-18.101$&$-18.506$&80&$2\pi\times14$&\\\hline
    $x$&-2&$-17.6375$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $z$&-5&$-18.1685$&$-18.506$&60&$2\pi\times14$&\\\hline
    $y$&-2&$-17.2825$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-6&$-18.101$&$-18.506$&80&$2\pi\times14$&\\\hline
    $x$&-1&$-18.03$&$-18.42$&25&$2\pi\times34$&F1\\\hline
    $z$&-5&$-18.1685$&$-18.506$&60&$2\pi\times14$&\\\hline
    $y$&-1&$-17.875$&$-18.47$&21&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 5}
This group is repeated 7 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-5&$-18.1685$&$-18.506$&90&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&38&$2\pi\times34$&F1\\\hline
    $z$&-4&$-18.236$&$-18.506$&75&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-5&$-18.1685$&$-18.506$&90&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&43&$2\pi\times34$&F1\\\hline
    $z$&-4&$-18.236$&$-18.506$&75&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&28&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 6}
This group is repeated 8 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-4&$-18.236$&$-18.506$&75&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&38&$2\pi\times34$&F1\\\hline
    $z$&-3&$-18.3035$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-4&$-18.236$&$-18.506$&75&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&43&$2\pi\times34$&F1\\\hline
    $z$&-3&$-18.3035$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&28&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 7}
This group is repeated 10 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-3&$-18.3035$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&38&$2\pi\times34$&F1\\\hline
    $z$&-2&$-18.371$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-3&$-18.3035$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&43&$2\pi\times34$&F1\\\hline
    $z$&-2&$-18.371$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&28&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 8}
This group is repeated 10 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-2&$-18.371$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&38&$2\pi\times34$&F1\\\hline
    $z$&-1&$-18.4385$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-2&$-18.371$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&43&$2\pi\times34$&F1\\\hline
    $z$&-1&$-18.4385$&$-18.506$&60&$2\pi\times17$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&28&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}
\subsubsection{Group 9}
This group is repeated 40 times.
\begin{center}
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Axis&Order&$\delta'$ (MHz)&$\delta'_0$ (MHz)&Time ($\mu$s)&Full Rabi frequency (kHz)&Power ramp\\\hline
    $z$&-1&$-18.3615$&$-18.429$&70&$2\pi\times9$&F2\\\hline
    $z$&-1&$-18.3615$&$-18.429$&100&$2\pi\times9$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&38&$2\pi\times34$&F1\\\hline
    $z$&-1&$-18.3615$&$-18.429$&70&$2\pi\times9$&F2\\\hline
    $z$&-1&$-18.3615$&$-18.429$&100&$2\pi\times9$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&25&$2\pi\times34$&F1\\\hline
    $z$&-1&$-18.3615$&$-18.429$&70&$2\pi\times9$&F2\\\hline
    $z$&-1&$-18.3615$&$-18.429$&100&$2\pi\times9$&F2\\\hline
    $x$&-1&$-18.03$&$-18.42$&43&$2\pi\times34$&F1\\\hline
    $z$&-1&$-18.3615$&$-18.429$&70&$2\pi\times9$&F2\\\hline
    $z$&-1&$-18.3615$&$-18.429$&100&$2\pi\times9$&F2\\\hline
    $y$&-1&$-17.875$&$-18.47$&28&$2\pi\times34$&F1\\\hline
  \end{tabular}
\end{center}

\bibliography{paper}
\end{document}